\vskip-0.6in Temporary Trade Barriers: When Will They End?

### 2017-09-18 draft.tex
remove un-needed material except for results, lit review and some of model

Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body. This paper sheds light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature. Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.

I show that the probability that the trade agreement is derailed increases in lobbying effort and decreases in the trade agreement tariff. Further, the lobby decreases its effort as the trade agreement tariff rises: it has less to lose the higher is the trade agreement tariff. When choosing the trade agreement tariffs, the executives therefore trade off the welfare they receive under the agreement with the probability that the agreement remains in force. Ultimately, I find that observed lobbying effort may often be low because the executives set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.

To see how this works, let's begin by considering the case of political certainty. I assume the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States. Lobbies, assumed for simplicity to represent import-competing interests only, would like to prevent or derail a trade agreement that would reduce their tariff protection. They work toward this end by exerting effort to convince their respective legislatures to vote against or ``break'' the trade agreement.\footnote{In this model, non-ratification is equivalent to breaking a trade agreement once in place because tariffs in a ``trade war'' are assumed to be set non-cooperatively just as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.}

For any trade-agreement tariff set by the executives, a lobby can calculate the effort level required to shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs. Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executives set the trade-agreement tariff so that lobbying is not worthwhile.

Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying.

The fact that zero contributions are predicted in equilibrium is not particularly satisfying. The model becomes more realistic, has smoother results and delivers additional intuition when I allow all actors to be uncertain about the weight the median legislator places on the profits of the lobbying industry. Similar to probabilistic voting models that incorporate policy motivation as in Roemer (1997), this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often so complex that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote. One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature. For instance, the lobbied legislators may promise to deliver the votes of their non-lobbied colleagues but may not be able to do so with certainty because those colleagues are subject to cross-pressuring from various sources.

One contribution of this model is that it shows how trade agreements can act as a domestic political commitment device in a broad set of environments. The executives set trade-agreement tariffs with an eye to discouraging lobbying activity and the accompanying probability of abrogation. In addition to the standard terms-of-trade internalization, the executive's agenda-setting power allows it to reduce tariffs until lobbying activity is at the optimal level relative to a non-cooperative setting, reducing the amount of political pressure faced by the legislatures and the resultant expected tariffs.

This is not a time inconsistency \`{a} la Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved by changing firms' investment decisions via commitment to a trade agreement. Instead, this is an inconsistency between the preferences of the executive and legislative branches. Although the legislatures have the final word, the executives use the trade agreement to reduce lobbying effort, muting the legislature's protectionist bias and bringing the outcome more in line with their relatively liberal preferences.

A further contribution is the demonstration that lobbying effort and the probability of trade disruptions vary in nuanced ways with the amount of political uncertainty present. Lobbying incentives are crucial but their impact is not necessarily as straightforward as predicted by models with unitary governments. Gawande, Krishna and Olarreaga (2009) provide cross-country evidence that measures such as the number of checks and balances on the power of the legislature have predictive power for the PFS ``welfare-mindedness'' parameter. Thus, although empirical tests are outside the scope of this paper, there is suggestive evidence that political uncertainty is indeed empirically important.

To the best of my knowledge, this is also the first attempt to model lobbying specifically aimed at derailing trade agreements. Although this kind of politically-driven failure is not rare, prior models have not allowed exploration of the endogenous dynamics behind them.

The addition of a rich structure of government, tailored carefully to the trade policy making process, also helps shed light on the empirical puzzle surrounding the PFS model. Empirical tests of the model\footnote{cfr. Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?

Attempts to more fully model the lobbying process have reduced the parameter estimates for the government's welfare considerations somewhat, but they still indicate that the U.S. government, for instance, values social welfare at least twenty times as much as contributions.\footnote{cfr. Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others. Gawande and Magee (2012) provide a summary of the extensive literature, while Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.} As suggested by Gawande and Hoekman (2006), in the non-unitary government model, these estimates and the stylized facts are not inconsistent. The ``missing contributions,'' can be explained as an equilibrium phenomenon arising from the agenda-setting power of the executive branch. In particular, in this model, there is no inconsistency for the 15$\%$ or so of sectors that receive protection but apparently put forth no lobbying effort. In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.

Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the legislatures' decision to abide by or break the trade agreement is modeled on the ``Fast Track / Trade Promotion Authority'' that the U.S. Congress periodically grants to the Executive branch.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries. The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for this dynamic to occur. Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether (Henisz and Mansfield 2006). In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways.

\subsection{Related Literature}
This work rests on the foundation of Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements. The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach views the ``Trade War'' as a crucial subgame. The model of preferential trade agreements (Grossman and Helpman (1995a)), although treating the government as a unitary actor, is closer to this approach.

@@

@@

@@

@@

\subsection{Information and Equilibrium Selection}
\label{sec:info}
I examine a simple class of equilibria that have three key features. First, information about political uncertainty is symmetric. However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature. Thus the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr. Grossman and Helpman (1995b), page 685.} Since players in the same country can take advantage of more information than those who are in different countries, the solution concept is perfect public equilibrium (PPE).

@@

\begin{result}
  When the trade agreement remains in force with positive probability, lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
  \label{res:lobby}
\end{result}

@@

We are now prepared to examine the executives' optimal choice of trade agreement tariffs. Because joint executive welfare is decreasing in trade agreement tariffs for $\bta$ above the executives' preferred tariffs, we have the following fundamental feature of the executives' problem:
\begin{lemma}
  The executives face the following trade off when choosing $\bta$: higher tariffs decrease the probability that the trade agreement will be broken but also decrease welfare when the agreement is in force.
  \label{res:to}
\end{lemma}
I show in the \hyperlink{int_soln}{Appendix} that it is never optimal in a symmetric equilibrium for the executives to choose $\bm{\tau^a}$ at or below the executives' most preferred level if the legislatures will have an opportunity to break the trade agreement. This is because there are only second-order losses from raising tariffs slightly from the most-preferred level yet there are first-order gains in reducing the probability that the agreement will be broken.

@@

\section{An Example}
\label{sec:example}
In this simple parameterization of the model, the fundamentals are chosen to match Bagwell and Staiger (2005): $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$. Home-country imports of $X$ and exports of $Y$ are $M_X(P_X)= 1 - \frac{3}{2}P_X$ and $E_Y(P_Y)= 2P_Y -1$. Market clearing implies that world and home prices of $X$ are $P_X^W = \frac{4-3\tau}{7}$ and $P_X = \frac{4+4\tau}{7}$. Foreign is taken to be symmetric.

\subsection{Trade War Tariffs}
The median legislator's welfare can be written as 
\[
  W_{\mathit{ML}}^X(\tau,\ga(e,\ve)) + W_{\mathit{ML}}^Y(\tau^*) = \left\{\frac{9}{98} - \frac{5}{49}\tau - \frac{34}{49}\tau^2 +\frac{1}{98}\ga(e,\ve)\left[ 8 + 16\tau + 8\tau^2 \right] \right\}+ \frac{25}{98} - \frac{3}{49}\tau^* + \frac{9}{49}(\tau^*)^2.
\]
%where $W_{\mathit{ML}}^X(\tau,\ga(e,\ve))$ is the utility derived from consumer surplus, producer surplus and tariff revenues in the import-competing industry and $W_{\mathit{ML}}^Y(\tau^*)$ is the utility derived from consumer and producer surplus in the exporting industry.

The Nash tariff that results from unilateral maximization is\footnote{The second order condition is satisfied for all realizations of $\ga < 17/2$. Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$, I will assume this condition is satisfied.}
\[
  \tau^n = \frac{8\ga(e,\ve)-5}{68-8\ga(e,\ve)}
\]

To predict the Nash tariff, the political weighting function and form of political uncertainty must be specified. Take $\ga(e,\ve) = 1.25 + e^{.2} + \ta$ with $\ve$ distributed uniformly on $[-0.25,0.25]$. Facing this specification of the political process, the lobby maximizes its objective function at $e_n = 0.00166$, producing an expected Nash tariff of $0.129$. Net profits are $0.10239$ in terms of the numeraire and lobbying effort is $1.60\%$ of gross profits.

\subsection{Break Decision}

\begin{comment}
\begin{figure}
\begin{center}
\includegraphics[height=2.50in, width=3in]{contributions.jpg}
\end{center}
\caption{Lobbying Effort\label{fig:cont}}
\end{figure}
%graph is created in lobby_fix.m
\end{comment}

\subsection{Trade Agreement}
The break probability never reaches zero for this parameterization so the executives' welfare function is concave everywhere. Assuming that each legislature has the opportunity to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers, they set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures at the trade maintenance phase of $0.0007$ and a total break probability of $0.505$. Total lobbying expenditure is $1.54\%$ of the expected profits of $0.98591$. The expected tariff is $0.103$.

We can compare this against different benchmarks. The most stark is the trade-war outcome, which is the outcome that would prevail in the absence of any trade agreement. The tariff in the executive-formed trade agreement is about 60$\%$ of the Nash tariff of $0.129$, and the expected tariff given the probability that the agreement will be broken is 80$\%$ of the Nash level. Lobbying expenditures in the agreement-maintenance phase are about 40$\%$ of the those in the Nash game, while expected expenditures in the executive-led trade-agreement scenario are 90$\%$ of what they would be if the legislatures set tariffs unilaterally. Although the welfare-maximizing governments here are not able to set and maintain tariffs at zero as they would like, they are able to achieve significant reductions in tariff levels through the use of the trade agreement.

Another benchmark is the scenario in which the legislatures make use of a trade agreement without the involvement of the executives. As in Bagwell and Staiger (2005), I find when maximizing their joint welfare in a legislature-led trade agreement, the cooperative tariff levels will be set at
\[
  \tau^L = \frac{4\ga(e,\ve)-4}{25-4\ga(e,\ve)}, \ \tau^{*L} = \frac{4\ga(e^*,\ve)-4}{25-4\ga(e^*,\ve)}
\]
The legislatures are able to use the agreement to internalize the terms of trade externality, making political influence more expensive for the lobbies. Lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case and $2.64\%$ of the gross profits of $0.10204$---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than in the executive-led trade agreement. Comparing net expected profits across the three scenarios, they are highest in the Nash game, lowest under the trade agreement and intermediate when the legislatures make policy directly.

\section{Discussion}
\label{sec:dis}

\subsection{Separation of Powers}
\label{sec:sep_powers}
We can isolate the results that derive from the assumption that power over trade policy is shared between the executive and legislative branches if we assume that there is no uncertainty at the break stage.

When $\ga(e_b)$ is deterministic, the lobby knows the precise contribution it must make for any given trade agreement tariffs $\bta$ to induce the legislature to break the agreement. Here, the lobby's contribution \textit{increases} in $\bta$ since the higher are the trade agreement tariffs, the larger is the political weight that is required to induce the legislature to find them unsatisfactorily low. As long as trade war profits net of the required contribution are greater than trade agreement profits, the lobby will induce a trade war; otherwise, it is not in the lobby's interest to make any contribution. Facing this behavior, the executives set the lowest trade agreement tariffs that make it prohibitively expensive for the lobby to have the agreement broken.

In the example of Section~\ref{sec:example}, if the executives set the trade agreement facing no uncertainty, the tariff level is $0.106$, the lobby exerts zero effort, and the agreement remain in force with probability 1. If instead the executives were to set trade-agreement tariffs of $0.105$, the lobby would contribute $0.0025$ and the legislature would break the agreement with probability 1. This demonstrates both the agenda-setting power of the executives and the stark discontinuities induced when political uncertainty is not an issue.

This case highlights the manipulation of tariffs to discourage lobbying and the disconnect that can arise between the tariffs that are chosen and the preferences of governmental actors. The executives here prefers zero tariffs while the legislators prefer $\tau^a = 0.05$ when $e_b = 0$. However, the trade agreement tariffs are set at $0.106$ precisely to ensure that $e_b =0$ so that we have a legislature that upholds the agreement.

\subsection{The Role of Political Uncertainty}
\label{sec:uncertainty}
Adding political uncertainty allows for positive lobbying on the equilibrium path, but this is not the only additional insight. For simplicity, I assume uncertainty is mean-zero. Lobbying behavior in a trade war is not altered by mean-zero uncertainty and so neither are expected trade-war tariffs. At earlier stages, uncertainty alters optimizing behavior by both the lobby and the executives. Consider adding a very small amount of uncertainty to the example of the previous section: take $\ve$ distributed uniformly on $[-0.01,0.01]$. Here, the executives find it optimal to set the tariffs to completely disengage the lobby, ensuring the agreement remains in force. However, the tariff level will be different than when $\ta_b=0$ because instead of making the lobbying decision according to the certainty condition $\pi(\tau^n) - e_n - \pi(\tau^a) > e_b$, the lobby makes this decision according to Condition~\ref{ine:lobint}. Relative to the example with $\ta_b=0$, the tariff can be reduced slightly to $0.105$.

\subsection{Relation to Grossman and Helpman's `Protection for Sale'}
\label{sec:gh}
I have introduced a legislative welfare function that admits a wide range of political processes within a non-unitary legislature as well as unitary policy-making when special interests face non-constant returns to lobbying activity. The non-linear relationship between lobbying effort and tariffs follows Dixit, Grossman and Helpman (1997) and is a departure from the `Protection for Sale' welfare function, which assumes that a unitary government maximizes the sum of contributions and some fraction, $a$, of social welfare.

Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a benefit is that it does not require returns to lobbying effort to be constant. The more flexible form of the weighting function also removes the equilibrium indeterminacies and inability to speak to distributional questions that are implied by the linear form and its transferable utility (Dixit, Grossman and Helpman 1997).

The elegant PFS form is able to capture key cross-industry predictions on import tariffs. A perhaps unintended byproduct in the subsequent literature seems to have arisen in the form of a focus on matching fine details of governments' welfare-mindedness using a model that has a very simple unitary government and is thus better suited for other purposes.

It is clear that trade policy is often shaped in a complicated process involving multiple actors (see, for instance, Destler (2005) for the U.S.). The model presented herein is a first attempt to more richly model the policy-making process and thus shed light on open questions surrounding how government preferences and the details of political institutions impact trade policy outcomes.

\begin{comment}
\section{Extensions}
\label{sec:ext}

This analysis has been undertaken with the assumption that the import-competing industry is distributed evenly across legislative districts. By varying the way in which those profits enter the welfare function of the median legislator (before lobbying takes place), I can provide results about how district composition affects trade policy.

There are also several directions to explore in comparing different government structures. I start by describing the value of a trade agreement in an institutional setting of divided government.
\begin{result}
  Lobbying effort is reduced by the presence of executive branches who form a trade agreement (where the counterfactual is legislatures setting unilateral trade policy).
\end{result}
\textcolor{blue}{Don't yet know what assumptions I need to make this work} \\

Comparing this model to the unitary-government model is a delicate task. The most straightforward approach from the current structure is to disregard the executives and allow the legislative branches to form a trade agreement cooperatively after lobbying takes place. Note that we can retain the stochastic nature of the tariff-setting, but there is no vehicle for the agreement to be broken. In this case, the trade agreement tariff in home, $\tau_U^a$ will maximize
\[
  W_X(\tau,\ga_E) + W_X^*(\tau) = \frac{9}{98} - \frac{5}{49}\tau - \frac{34}{49}\tau^2 +\frac{1}{98}\ga(e,\ve)\left[ 8 + 16\tau + 8\tau^2 \right] + \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2 
\]
and will thus be set at
\[
  \tau_U^a = \frac{4\ga(e,\ve)-4}{25-4\ga(e,\ve)}.
\]
$\tau_U^a < \tau^n$ for all $\ga(e,\ve) < 7/4$, while $\ga(e,\ve)=7/4$ results in the prohibitive tariff of $1/6$. This is the standard result that a trade agreement internalized the terms of trade externality and results in a lower tariff than unilateral policy making.

\textcolor{blue}{will an executive inserted into this situation set higher or lower tariffs, what will happen to $e$?}

What if we add executive branches with essentially the same preferences as the legislature? That is, set $\ga_E = \expect[\ga(e,\ve)]$. \textcolor{blue}{It's hard to say how $e$ should be chosen here...still need to figure that out. But if we can get around the weird time inconsistency problem...} What agreement tariffs will the executives choose? It's easy to see that they will not choose $\tau^a < \tau_U^a$, as $\tau_U^a$ is their ideal point if there were no possibility of the agreement being broken, and lower tariffs only incentivize more lobbying and a higher probability that the agreement will be broken.

Thus, if an interior solution exists \textcolor{blue}{[should have conditions to insert here]}, the executive will exploit the same trade-off explored earlier, raising $\tau^a$ a bit to reduce the probability that the agreement will be abrogated. So interestingly, we can see that executives with the same political preferences as the legislature will choose higher trade agreement tariffs than the legislature would by itself due to the possibility that the agreement can be broken and the impact of the tariff choice on lobbying. We will also therefore have a reduced level of contributions, and if a trade war were possible, lower trade war tariffs because of it.

Along these lines, allowing for asymmetries in the preferences and political processes between the countries will provide important insights as well.

I also plan to use the richness of this model to parse out how much of the changes in tariffs comes from the fact that the executives place lower weight on producer profits and how much comes from the fact that they make an agreement---that is, compare the effects of structure versus preferences.

\subsection{Asymmetric Trade Agreements}
\label{sec:asym}
The assumption that the legislatures do not have the opportunity to break a trade agreement simultaneously serves to simplify the analysis of the interaction between the lobbies and the consequences of relaxing it will be discussed in the following section. Maintaining it for now, however, facilitates examination of the assumption that trade agreements must be symmetric. This may be quite realistic in some settings where additional political constraints would make ``unfair'' arrangements unpalatable. However, it is important to point out that, for some parameter choices, the joint welfare of the executives is in fact maximized by an asymmetric agreement.

The parameterization from the example in Section~\ref{sec:example} is one such instance. The optimal unrestricted trade agreement is to set tariffs in one country (say home) at $0.062$ and those in the other (foreign) at 0. In this case, the equilibrium break probability in home, if the home legislature is afforded the opportunity, is zero, while the foreign legislature will break the agreement with probability 1 if given the chance. With the assumption from the example that each legislature gets the opportunity to repudiate the agreement half the time, this results in the trade agreement being broken with probability $0.50$, just less than the $0.505$ of the optimal symmetric agreement with both tariffs set at $0.078$. Because both the trade agreement tariffs and the expected chance that the agreement will remain in force are lower, expected welfare for the executives is higher.

What is behind the rather surprising result that a lower break probability can be achieved when both tariffs are strictly lower than in the symmetric agreement? Here, the executives relinquish any chance of passage of the trade agreement in the foreign country (in essence hoping that its legislature will not have a chance to act). Once they decide to pursue this course, it is optimal to reduce $\tau^{*a}$ to zero. This drastically reduces the home legislature's incentive to break the trade agreement. Because of the large direct impact of $\tau^{*a}$ in reducing the break probability, the home tariff is not required to make such a large contribution and can be lowered, which in turn reduces the lobby's incentive to exert effort (see Result~\ref{res:lobby}). This effect is strong enough that the executives can set a tariff level lower than that of the symmetric agreement that is sufficient to disengage the home lobby (see Inequality~\ref{ine:lobint}).

It is useful to keep in mind that this result occurs in a completely symmetric environment. Recall that, in general, the executives maximize joint welfare in Equation~\ref{eq:jv2}. If the chance that the legislatures will have the opportunity to break the agreement ($o$ and $o^*$) are not equal, or if another significant aspect of the economic or political situation is asymmetric, the importance of this result grows in a potentially substantial way. In particular, if only one country faces a legislative constraint, an asymmetric agreement seems highly likely as in the large body of literature investigating the Schelling conjecture.

\subsection{Break Opportunities in Both Countries}
\label{sec:two}
Although it changes their expected payoffs, extending the model to allow both legislatures the opportunity to break the trade agreement simultaneously does not impact the legislatures' incentives as each decision to break or not break the agreement is independent of the other.

However, the lobbies' incentives are altered because the total probability of a break with which they weight profits changes. Now, in addition to their own effort positively influencing their own legislature to break the agreement, the probability of a break in the trade agreement also increases in the effort of the other country's lobby. As might be expected, the lobbying effort in each country turns out to be a negative function of that in the trading partner; that is, to some extent, the lobbies have an incentive to free ride.

Without making stronger assumptions, it is not possible to say how much the effort of each lobby will be reduced therefore whether total lobbying effort rises or falls. What is clear is that the optimization of the executives becomes significantly more complex as they are now faced with the problem of how to best exploit the free-riding dynamics that occur between the lobbies. 
\end{comment}

\section{Conclusion}
\label{sec:concl}
I have shown that the legislature both breaks trade agreements with a higher probability and sets higher trade war tariffs when lobbying activity increases, while the probability with which it breaks agreements decreases in the domestic trade agreement tariff. Because the lobby decreases its effort in response to higher trade agreement tariffs, the executives face a trade-off between the welfare derived while a trade agreement is in force and the probability with which the agreement actually enters into force.

I have also shown that in a policy-making structure in which power is shared, a less politically-motivated executive can utilize an international trade agreement to reduce the political pressure on the ratifying body and therefore increase the probability that the agreement will remain in force. Thus, in a model with a richer description of government structure, a broad political-commitment role for trade agreements can arise.

The executives' incentive to raise tariffs in order to reduce lobbying effort provides insight into the empirical puzzle in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare. The intuition is clear: the observed lobbying expenditure levels may in fact be low \textit{because} tariffs have been raised sufficiently high to prevent political pressure and the increased risk of a costly trade disruption it engenders.

That lobbying and tariff levels are related in systematic ways to the amount of political uncertainty present suggests interesting avenues for future empirical work. Several directions for future theoretical work also seem potentially fruitful, including removing the assumption of perfect enforceability and supporting cooperation through repeated interaction and generalizing the model to the case of multiple lobbies.

%\begin{comment}
\section{Appendix}
\label{sec:appendix}
\noindent \textbf{\hypertarget{Pr_bincC}{Proof of Result~\ref{res:bincC}}}:\\
Substituting from Equation~\ref{eq:ml}, Equation~\ref{eq:b} can be re-written as
\begin{multline}  
  b(e_b,\bta,\btn) = \Pr [ \mathit{CS}_X(\tau^n) + \ga(e_b,\ve_b) \cdot \pi_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \pi_Y(\tau^{*n}) + \mathit{TR}(\tau^n) >  \\ \mathit{CS}_X(\tau^a) + \ga(e_b,\ve_b) \cdot \pi_X(\tau^a) + \mathit{CS}_Y(\tau^{*a}) + \pi_Y(\tau^{*a}) + \mathit{TR}(\tau^a) ]
\end{multline}
Rearranging, we have $b(e_b,\bta,\btn) = $
\begin{multline}  
  \textstyle \Pr \Big[ \frac{\mathit{CS}_X(\tau^n) + \pi_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \pi_Y(\tau^{*n}) + \mathit{TR}(\tau^n)  -\mathit{CS}_X(\tau^a) - \pi_X(\tau^a) - \mathit{CS}_Y(\tau^{*a}) - \pi_Y(\tau^{*a}) - \mathit{TR}(\tau^a)}{\pi_X(\tau^a) - \pi_X(\tau^n)} + 1 \\ \textstyle < \ga(e_b,\ve_b) \Big]
  \label{eq:b_ex}
\end{multline}
%\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. By Assumption~\ref{as:ga_c}, the right side of the inequality is increasing and concave in $e_b$. Thus $b(e_b,\bta,\btn)$ is increasing and concave in $e_b$.    $\hfill\blacksquare$
\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. Call it $Z$. Thus we have $b(e_b,\bta,\btn) = \Pr [Z < \ga(e_b,\ve_b) ] = 1 - F_{\ga}(\ga=Z)$ where $F_{\ga}(\ga=Z) = F_{\ta}(\ta=h^{-1}(\ga,e))$ by the Change of Variables Theorem and Assumption~\ref{as:ga_ta} with $\ga = h(e,\ta)$ giving the change of variable.

Then $\frac{\partial b}{\partial e} = -\frac{\partial F_{\ta}(\ta=h^{-1}(\ga,e))}{\partial \ta}\frac{\partial h^{-1}(\ga,e)}{\partial e} = -f_\ta(\ta) \frac{\partial h^{-1}(\ga,e)}{\partial e}$ and $\frac{\partial^2 b}{\partial e^2} = -f_\ta(\ta) \frac{\partial^2 h^{-1}(\ga,e)}{\partial e^2}$. Because $\ga = h(e,\ta)$ is increasing and concave in $e \ \forall \ta$ by Assumption~\ref{as:ga_c}, its inverse is decreasing and convex $\left(\frac{\partial h^{-1}(\ga,e)}{\partial e}\leq 0; \ \frac{\partial^2 h^{-1}(\ga,e)}{\partial e^2} \geq 0 \right)$ The pdf of $\ta$ is non-negative, so $\frac{\partial b}{\partial e} \geq 0$ and $\frac{\partial^2 b}{\partial e^2} \leq 0$. $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_leg_astar}{Proof of Lemma~\ref{lem:leg_astar}}}:\\
It must be shown that the left hand side of the inequality in Expression~\ref{eq:b_ex} is decreasing in $\tau^{*a}$. The derivative of this quantity with respect to $\tau^{*a}$ is 
\begin{equation}
  \frac{-\left(\pi_X(\tau^a) - \pi_X(\tau^n)\right)\left(\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \pi_Y(\tau^{*a})}{\partial \tau^{*a}}\right)}{\left(\pi_X(\tau^a) - \pi_X(\tau^n)\right)^2} = \frac{\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \pi_Y(\tau^{*a})}{\partial \tau^{*a}}}{\pi_X(\tau^n) - \pi_X(\tau^a)}.
  \label{appex:astar}
\end{equation}

\noindent The price of good $Y$ is decreasing in $\tau^{*a}$, so consumer surplus is increasing and profits are decreasing in $\tau^{*a}$. Because $Y$ is being exported, the decrease in profits is larger than the increase in consumer surplus, making the numerator negative. Since profits in the denominator are increasing in $\tau$, the expression in Equation~\ref{appex:astar} is negative for all $\tau^{*a}$.  $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_leg_a}{Proof of Lemma~\ref{res:leg_a}}}:\\
Using the logic of the proof of Lemma~\ref{lem:leg_astar}, the effect on the break probability is determined by the sign of the derivative of the left hand side of the inequality in Expression~\ref{eq:b_ex} with respect to $\tau^a$; to show that the break probability is decreasing in $\tau^a$, I must demonstrate that this derivative is positive. Labeling the numerator of that expression $\left[ W(\btn)-W(\bta)\right]$ (for the change in social welfare), this derivative can be written
\begin{equation}
  \frac{\left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \pi_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \pi_X(\tau^a)}{\partial \tau^a}}{\left(\pi_X(\tau^a) - \pi_X(\tau^n)\right)^2}.
  \label{appex:a}
\end{equation}

\noindent $\left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)$ is always positive by Assumption~\ref{as:ga_l_e}. Because the optimal unilateral tariff for large welfare-maximizing governments is positive (call it $\tau^O$), $\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \pi_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right)$ is increasing up to $\tau^O$ and decreasing above it. Thus the first summand is increasing up until $\tau^O$ and decreasing thereafter.

Because total social welfare is maximized at $\tau^a = \tau^{*a} = 0$, $W(\btn)-W(\bta)$ is always negative, whereas profits are increasing in $\tau^a$, so the second summand is positive everywhere. With a positive denominator, we thus have that the derivative is positive on $[0,\tau^O]$.

It is also positive over the remaining $(\tau^O,\tau^n)$. To see this, add $\left(\tilde{\Gamma} - 1 \right) \frac{\partial \pi_X(\tau^a)}{\partial \tau^a} \left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)$ to the first summand and subtract it from the second. For any particular value of $\tilde{\tau}^a$, one can choose the $\tilde{\Gamma}$ weight that would make $\tilde{\tau}^a$ the preferred unilateral tariff; this makes the derivative in the first summand zero. Having subtracted the same quantity from the second summand modifies the welfare difference in the second summand to be maximized at $\tilde{\tau}^a$ so that this term is always negative, thus ensuring the result.  $\hfill\blacksquare$

>>>>

\subsection{Related Literature}
\textbf{WILL NEED MAJOR OVERHAUL}

This work rests on the foundation of Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements. The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach views the ``Trade War'' as a crucial subgame. The model of preferential trade agreements (Grossman and Helpman (1995a)), although treating the government as a unitary actor, is closer to this approach.

@@

\textbf{REVISE GOVERNMENT STRUCTURE}

@@

\textbf{LOBBY'S OBJECTIVE IS DIFFERENT HERE}

@@

\textbf{NOT SURE I KNOW EXACTLY WHAT THE GOVERNMENT'S OBJECTIVE IS, DIVIDED GOV'T EVEN?}

@@

\subsection{Information and Equilibrium Selection}
\label{sec:info}

\textbf{NOT SURE AT ALL ABOUT THESE FEATURES; it's not a two-country model}

I examine a simple class of equilibria that have three key features. First, information about political uncertainty is symmetric. However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature. Thus the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr. Grossman and Helpman (1995b), page 685.} Since players in the same country can take advantage of more information than those who are in different countries, the solution concept is perfect public equilibrium (PPE).

@@

\begin{result}
  When the trade agreement remains in force with positive probability, lobbying effort is weakly decreasing in trade agreement tariffs.
  \label{res:lobby}
\end{result}

@@

We are now prepared to examine the executives' optimal choice of trade agreement tariffs. Because joint executive welfare is decreasing in trade agreement tariffs for $\bta$ above the executives' preferred tariffs, we have the following fundamental feature of the executives' problem:
\begin{lemma}
  The executives face the following trade off when choosing $\bta$: higher tariffs decrease the probability that the trade agreement will be broken but also decrease welfare when the agreement is in force.
  \label{res:to}
\end{lemma}

@@

\section{Conclusion}
\label{sec:concl}

\section{Appendix}
\label{sec:appendix}
\noindent \textbf{\hypertarget{Pr_bincC}{Proof of Result~\ref{res:bincC}}}:\\

\section{References}
